---
title: "Introduction to data"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: ./lab.css
---


```{r load-data, eval=TRUE}
source("more/cdc.R")
```


```{r names, eval=TRUE}
names(cdc)
```

This returns the names `genhlth`, `exerany`, `hlthplan`, `smoke100`, `height`, 
`weight`, `wtdesire`, `age`, and `gender`. Each one of these variables 
corresponds to a question that was asked in the survey.  For example, for 
`genhlth`, respondents were asked to evaluate their general health, responding
either excellent, very good, good, fair or poor. The `exerany` variable 
indicates whether the respondent exercised in the past month (1) or did not (0).
Likewise, `hlthplan` indicates whether the respondent had some form of health 
coverage (1) or did not (0). The `smoke100` variable indicates whether the 
respondent had smoked at least 100 cigarettes in her lifetime. The other 
variables record the respondent's `height` in inches, `weight` in pounds as well
as their desired weight, `wtdesire`, `age` in years, and `gender`.

1.  How many cases are there in this data set?  How many variables?  For each 
    variable, identify its data type (e.g. categorical, discrete).

## Return dimensions(rows and columns), use the str command, define each column's data type

```{r head, eval=TRUE}
dim(cdc)
str(cdc)
```


```{r, echo=FALSE}
table_3<-read.table(text="
Column_name Numerical/Categorical Discrete/Coninuous/Ordinal
exerany Categorical NA
hlthplan Categorical NA 
smoke100 Categorical NA
height Numerical Discrete
weight  Numerical Discrete
wtdesire Numerical Discrete
age Numerical Discrete
gender categorical NA
wdiff Numerical Discrete
" ,  header=TRUE, stringsAsFactors=FALSE)
print (table_3)

```



2.  Create a numerical summary for `height` and `age`, and compute the 
    interquartile range for each. Compute the relative frequency distribution for
    `gender` and `exerany`. How many males are in the sample? What proportion of
    the sample reports being in excellent health?



## Create an interquartile function and print out several frequency distribution functions

```{r table-smoke-gender, eval=TRUE}
summary(cdc$height)
summary(cdc$age)

#Create IQR FUNCTION
quartile_range <- function(x){sorted <- sort(x)
    lower=median(sorted[1:10000])
    upper=median(sorted[10001:20000])
    inter= upper-lower
    return (c("1st_quart"=lower,"3rd_quart"=upper,"inter_quart"=inter))

}

#CREATE FREQUENCY TABLES
exercise_freq_dist <- table(cdc$exerany)/20000*100
gender_freq_dist <-   table(cdc$gender)/20000*100
healt_freq_dist <-    table(cdc$genhlth)/20000*100
males_in_sample <- gender_freq_dist[1]*200
great_health <- healt_freq_dist[1] *200

# USE PASTE FUNCTION FOR PRINT STATEMENTS
heights <- (paste(c("height_1st_quart is","height_3rd_quart is","height_inter_quartile range is"),c(quartile_range(cdc$height))))
ages <- (paste(c("age_1st_quart is","age_3rd_quart is","age_inter_quartile range is"),c(quartile_range(cdc$age))))
genders <- paste(c("percent males is","percent females is"),c(gender_freq_dist))
exercised <-paste(c("percent whom didn't exercise","percent whom did exercise"),c(exercise_freq_dist)) 
sampled_males <- paste("number of males",males_in_sample)
excellent_reported_health <- paste("number of respondents reporting excellent health",great_health)


print (c(heights,ages,genders,exercised,sampled_males,excellent_reported_health))
```

```{r mosaic-smoke-gender, eval=TRUE}
mosaicplot(table(cdc$gender,cdc$smoke100))
```

+  Males are more Likely to smoke than females.  



3.  Create a new object called `under23_and_smoke` that contains all observations 
    of respondents under the age of 23 that have smoked 100 cigarettes in their 
    lifetime. Write the command you used to create the new object as the answer
    to this exercise.
    
## Subset of young smokers and a frequency distribution for that young smoker subset
+ How do smokers over 55 compare?
```{r}
#CREATE SUBSET DATA AND PASTE FOR PRINT
under23_and_smoke <- subset(cdc, age<23,smoke100)
younger_smokers <- paste(c("distribution of  people under 23 that don't smoke ","distribution of people under 23 that do smoke"),c(table(under23_and_smoke)/dim(under23_and_smoke)[1]))
over_55_smokers <- subset(cdc, age>55,smoke100)
older_smokers <- paste(c("distribution of  people over 55 that don't smoke ","distribution of  people over 55 that do smoke"),c(table(over_55_smokers)/dim(over_55_smokers)[1]))


print(c(younger_smokers, older_smokers))
```

+ Only 2 of every 5 young Americans smoke where as over 55% of adults over 55 smoke  

## Quantitative data


```{r boxplot-height, eval=TRUE}
boxplot(cdc$height)
```

You can compare the locations of the components of the box by examining the 
summary statistics.

```{r summary-height, eval=TRUE}
summary(cdc$height)
```



The notation here is new. The `~` character can be read *versus* or 
*as a function of*. So we're asking R to give us a box plots of heights where 
the groups are defined by gender. 

Next let's consider a new variable that doesn't show up directly in this data 
set: Body Mass Index (BMI) 
([http://en.wikipedia.org/wiki/Body_mass_index](http://en.wikipedia.org/wiki/Body_mass_index)). 
BMI is a weight to height ratio and can be calculated as:

\[ BMI = \frac{weight~(lb)}{height~(in)^2} * 703 \]

703 is the approximate conversion factor to change units from metric (meters and 
kilograms) to imperial (inches and pounds).

The following two lines first make a new object called `bmi` and then creates 
box plots of these values, defining groups by the variable `cdc$genhlth`.

```{r boxplot-bmi, eval=TRUE}
bmi <- (cdc$weight / cdc$height^2) * 703
boxplot(bmi ~ cdc$genhlth)
hist(bmi)

```



4.  What does this box plot show? Pick another categorical variable from the 
    data set and see how it relates to BMI. List the variable you chose, why you
    might think it would have a relationship to BMI,  and indicate what the 
    figure seems to suggest. 
    
+ This shows that lower bmi is indicative of stronger health outcomes.  
    + Excellent health outcomes fall in a much lower and narrower bmi range.  
    + It appears positive health outcomes and bmi are negatively correlated
+ the data does appear to have alot of outliers, however it is a large sample. 

## Boxplot of bmi categorized by exercise column
+ Observations from box plot plotted below
    
    + I would assume people who exercise have lower BMI.  
    + It appears the IQR of exercisers BMI is lower as well   
    + There appears to be a decent amount of repondents outside the interquartile range in both exercisors and non exercisors
```{r}
boxplot(bmi ~ cdc$exerany)
summary(cdc$exerany==1)
```





```{r hist-bmi, eval=TRUE}
hist(bmi, breaks = 50)
```


## On Your Own

```{r}
library(stringr)
library(XML)
library(maps)
library(data.table)
library(dplyr)
library(psych)
library(sjmisc)
library(ggplot2) 
```

5. Make a scatter plot of weight versus desired weight. Describe the 
    relationship between these two variables.
    
```{r}
library(ggplot2)
ggplot(cdc, aes(x = cdc$weight, y = cdc$wtdesire)) +
    theme_bw() +
    geom_point()
```

+ There is a small positive correlation between actual weight and expected weight.  
+ But it appears that this correlation changes depending on body weight.   
    + At around 100 pounds, most people want to be heavier.  
    + At around 200 pounds most people want to be thinner.  
+ There are some outliers that likely need to be removed as noone really wants to be over 600 pounds.  
    + These people are either providing misleading information or there was some clerical error collecting the data.   

## Eliminate outlier under the assumption that noone wants to weigh over 500 pounds
+ getting rid of the outlier also makes for a much clearer graph of the relationship between our variables 

6. Let's consider a new variable: the difference between desired weight 
    (`wtdesire`) and current weight (`weight`). Create this new variable by 
    subtracting the two columns in the data frame and assigning them to a new 
    object called `wdiff`.
```{r}
#CREATE WDIFF COLUMN AND PRINT PLOT OF DATA WITHOUT OUTLIER
cdc$wdiff <-cdc$weight-cdc$wtdesire
sub_set_cdc <- subset(cdc,cdc$wdiff> -300)
library(ggplot2)
ggplot(sub_set_cdc, aes(x = sub_set_cdc$weight, y = sub_set_cdc$wtdesire)) +
    theme_bw() +
    geom_point()

```

7.  What type of data is `wdiff`? If an observation `wdiff` is 0, what does 
    this mean about the person's weight and desired weight. What if `wdiff` is 
    positive or negative?
    
+ Wdiff is integer column.  It represents discrete Numerical data
+ When 0 it means a person desired weight is equal to their actual weight
+ When positive it means the person wants to weigh less than they weigh
+ When negative it means the person wants to weigh more than they weigh

## Some visual display and summary of our new Wdiff column
```{r}

hist(sub_set_cdc$wdiff,breaks=41)
summary(sub_set_cdc$wdiff)
summary(cdc$wdiff)

d <- density(cdc$wdiff)
plot(d, main="Kernel Density of desired weight")
polygon(d, col="red", border="blue")

```

8.    Describe the distribution of `wdiff` in terms of its center, shape, and 
    spread, including any plots you use. What does this tell us about how people 
    feel about their current weight?

+ Our histogram is rightward skewed with an interquartilerange of 21, and a spread of over 410. The range of the original data, with outliers, would be even larger at around 800.  
+ The rightward skew implies that there are more people that want to lose larger amounts of weight then want to gain larger amounts of weight
+ Most people,50% of our data set, want to remain the same weight or lose up to 21 pounds.
+ Only 25% of our data set wants to gain any weight at all


9.   Using numerical summaries and a side-by-side box plot, determine if men tend
    to view their weight differently than women.

## WDIFF plotted side by side by gender
+ Red= Men , Blue= Women
```{r}
# SUBSET WDIFF COLUMN BY GENDER, PLOT GENDERS SIDE BY SIDE
sub_set_cdc_women <-subset(sub_set_cdc,sub_set_cdc$gender=="f") 
sub_set_cdc_men <-subset(sub_set_cdc,sub_set_cdc$gender=="m") 
summary(sub_set_cdc_women$wdiff)
summary(sub_set_cdc_men$wdiff)
gender_df <- rbind(data.frame(fill="blue", obs=sub_set_cdc_women$wdiff),
            data.frame(fill="red", obs=sub_set_cdc_men$wdiff))
ggplot(gender_df, aes(x=obs, fill=fill)) +
  geom_histogram(binwidth=10, colour="black", position="dodge") +
  scale_fill_identity()
```



+ While some men think they should gain weight, fewer women feel as though they should gain weight
+ A large amount of men and women think their weight is fine, although more men are comfortable with their current weight  
+ In almost all ranges, more women feel like they should lose weight then men feel like they should lose weight

10. -   Now it's time to get creative. Find the mean and standard deviation of 
    `weight` and determine what proportion of the weights are within one 
    standard deviation of the mean.
```{r}
# GET MEAN, GET STD DEVIATION, CREATE A SUBSET WITHIN 1 STD DEVIATION IN EACH DIRECTION
mean_cdc <- mean(sub_set_cdc$weight)
sd_cdc <- sd(sub_set_cdc$weight)
stand_dev_upper <-mean_cdc+sd_cdc
stand_dev_lower <- mean_cdc-sd_cdc 
cdc_data_within_sd <- subset(sub_set_cdc,sub_set_cdc$weight<stand_dev_upper &sub_set_cdc$weight>stand_dev_lower )
proportion_within_sd <- dim(cdc_data_within_sd)[1]/20000
print(paste(" the proportion of weights that fall within one standard deviation is",proportion_within_sd))
```




<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel 
from a lab written by Mark Hansen of UCLA Statistics.
</div>
